---
title: "NYPD Shootings Data"
author: "Joseph Bae"
date: "2023-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First import the tidyverse package which we'll be using extensively throughout this analysis.

```{r}
library(tidyverse)
```

Read in our data which is the historical NYPD Shootings Data available on the city of New York's website.

```{r}
ny_data = read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD")
```

Adjust column date types to make them easier to analyze, and also remove geolocation columns which I'm not interested in for this particular analysis. Finally, add a Year and Month column based on OCCUR_DATE.

```{r}
ny_data = ny_data %>% mutate(OCCUR_DATE = mdy(OCCUR_DATE))
ny_data = ny_data %>% mutate(BORO = as_factor(BORO))
ny_data = ny_data %>% mutate(LOC_OF_OCCUR_DESC = as_factor(LOC_OF_OCCUR_DESC))
ny_data = ny_data %>% mutate(PRECINCT = as_factor(PRECINCT))
ny_data = ny_data %>% mutate(JURISDICTION_CODE = as_factor(JURISDICTION_CODE))
ny_data = ny_data %>% mutate(LOC_CLASSFCTN_DESC = as_factor(LOC_CLASSFCTN_DESC))
ny_data = ny_data %>% mutate(LOCATION_DESC = as_factor(LOCATION_DESC))
ny_data = ny_data %>% mutate(STATISTICAL_MURDER_FLAG = as_factor(STATISTICAL_MURDER_FLAG))
ny_data = ny_data %>% mutate(PERP_AGE_GROUP = as_factor(PERP_AGE_GROUP))
ny_data = ny_data %>% mutate(PERP_SEX = as_factor(PERP_SEX))
ny_data = ny_data %>% mutate(PERP_RACE = as_factor(PERP_RACE))
ny_data = ny_data %>% mutate(VIC_AGE_GROUP = as_factor(VIC_AGE_GROUP))
ny_data = ny_data %>% mutate(VIC_SEX = as_factor(VIC_SEX))
ny_data = ny_data %>% mutate(VIC_RACE = as_factor(VIC_RACE))

ny_data = ny_data %>% select(-c(X_COORD_CD, Y_COORD_CD, Latitude, Longitude, Lon_Lat))

ny_data = ny_data %>% mutate(YEAR = as.integer(format(OCCUR_DATE, "%Y")))
ny_data = ny_data %>% mutate(MONTH = as.integer(format(OCCUR_DATE, "%m")))

summary(ny_data)
```

Let's see the trend of shooting incidents by Year, and also by Month.

```{r}
by_year = ny_data %>% group_by(YEAR) %>% summarize(Incidents = n())
by_month = ny_data %>% group_by(YEAR, MONTH) %>% summarize(Incidents = n())

by_year

ggplot(by_year) + geom_point(mapping = aes(x = YEAR, y = Incidents)) + geom_line(mapping = aes(x = YEAR, y = Incidents)) + labs(title = "NY Shooting Incidents by Year")

ggplot(by_month) + geom_col(mapping = aes(x = MONTH, y = Incidents, fill = Incidents)) + labs(title = "NY Shooting Incidents by Year and Month") + scale_x_continuous(breaks = c(2,4,6,8,10,12), labels = c(2,4,6,8,10,12)) + facet_wrap(~YEAR)

by_month %>% filter(Incidents == max(Incidents)) %>% select(YEAR, MONTH)
```

Distribution of incidents by borough

```{r}
ggplot(ny_data) + geom_bar(mapping = aes(x = BORO)) + labs(title = "NY Shooting Incidents by Borough")
```

Distribution of victim age, race, sex

```{r}
ggplot(ny_data) + geom_bar(mapping = aes(x = VIC_AGE_GROUP)) + labs(title = "NY Shooting Incidents by Victim Age Group")

ggplot(ny_data) + geom_bar(mapping = aes(x = VIC_RACE)) + coord_flip() + labs(title = "NY Shooting Incidents by Victim Race")

ggplot(ny_data %>% filter(!VIC_RACE == 'UNKNOWN')) + geom_bar(mapping = aes(x = BORO)) + coord_flip() + facet_wrap(~ VIC_RACE)

by_sex <- ny_data %>% group_by(VIC_SEX) %>% summarize(Incidents = n())
ggplot(by_sex, aes(x = "", y = Incidents, fill = VIC_SEX)) + geom_col() + coord_polar(theta = "y") + labs(title="Count of New York Shooting Incidents by Victim's Sex")
```

Distribution of perpetrator age, race, sex

```{r}
# 14,177 rows removed by filter
ggplot(ny_data %>% filter(PERP_AGE_GROUP %in% c("25-44", "18-24", "45-64", "<18", "65+"))) + geom_bar(mapping = aes(x = PERP_AGE_GROUP)) + labs(title = "NY Shooting Incidents by Perpetrator Age Group")

# 11,786 rows removed by filter
ggplot(ny_data %>% filter(!is.na(PERP_RACE)) %>% filter(!PERP_RACE %in% c("(null)", "UNKNOWN"))) + geom_bar(mapping = aes(x = PERP_RACE)) + coord_flip() + labs(title = "NY Shooting Incidents by Perpetrator Race")

# 11,449 rows removed by filter
by_sex <- ny_data %>% filter(PERP_SEX %in% c("M", "F")) %>% group_by(PERP_SEX) %>% summarize(Incidents = n())
ggplot(by_sex, aes(x = "", y = Incidents, fill = PERP_SEX)) + geom_col() + coord_polar(theta = "y") + labs(title="Count of New York Shooting Incidents by Perpetrator's Sex")
```

Modeling shooting incidents based on Month:

```{r}
by_month$MONTH2 = by_month$MONTH^2
qm = lm(data = by_month, Incidents ~ MONTH + MONTH2)
summary(qm)

x = 1:12
y = predict(qm, list(MONTH = x, MONTH2 = x^2))
df = data.frame(x, y)
ggplot(by_month) + geom_point(mapping = aes(x = MONTH, y = Incidents)) + geom_line(data = df, mapping = aes(x = x, y = y)) + scale_x_continuous(breaks = 1:12, labels = 1:12)
```